---
title: "ETC5521 Assignment 1"
subtitle: "Your title"
team: [emu]
author:
  - [FILL]
  - Min Min Soh
date: "`r Sys.Date()`"
output: 
  html_document
---

[This assignment is for ETC5521 Assignment 1 by Team `r rmarkdown::metadata$team` comprising of `r knitr::combine_words(rmarkdown::metadata$author)`.]{style="color:#006DAE;"}

```{r}
library(tidyverse)
```

# Introduction and motivation (Emily)

Global fishing has received a great deal of attention in the media for the past decades. The rise of world population over the decades have increased the demand for seafood consumption across the world. Coastal countries such as Iceland and Maldives still mainly rely on fishing as one of the most important food sources.  In this analysis, our first research question would learn about the seafood consumption of each country across the years. We will also study about how much seafood are being produced by each country overtime.  

Meanwhile, the rise of seafood demands has increased pressure on fish stocks across the world. The health of fish population in the world remains as a concern with the rise of global fishing. There are claims that ongoing erosion of aqua diversity will eventually lead to global collapse of fisheries by 2048. It then prompts us to have a closer look at how much of fish stocks are caught within the sustainable levels over the years without overexploiting the fish populations.

Fish farming, also being known as aquaculture come into place, help to contribute to the seafood production  while alleviating the pressure of wild fisheries. This inspires us to further analyse the change in the trend of aquaculture as compared to wild fish catch over time. 

We begin by describing the data in the next section, how we source it and how we prepare the data for analysis. In the analysis section, we present our observations through graphical displays. Our main tool is R, a programming language for statistical computing and graphics, which we use to clean, analyse and graph the data.  


# Data description 

[FILL] Source of the data, and detail the structure of the data (Rohan)

---- source --- worlddata ... (more sources)
---- table (what kind of data) - a short description of the variable class 
--- 

```{r}
#read data
farmed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/aquaculture-farmed-fish-production.csv')
captured_vs_farmed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/capture-fisheries-vs-aquaculture.csv')
captured <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/capture-fishery-production.csv')
consumption <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/fish-and-seafood-consumption-per-capita.csv')
stock <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/fish-stocks-within-sustainable-levels.csv')
fishery <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/global-fishery-catch-by-sector.csv')
production <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/seafood-and-fish-production-thousand-tonnes.csv')

```

# Questions of interest

[FILL] What questions might be possible to answer with this data, aiming for about 20. Bold the 3-4 questions you will address in the analysis

Q1 What is the contribution of each production sector in global fishery from 1950 ? 

Q2 ** What is the contribution of each country in the global fishery sector ? ** 

Q3 What is the share of type of fishes produced in each country ? 

Q4 What is the production level of each country by capturing over the years ?

Q5 What is the production level of each country by farming over the years  ?

Q6 ** What has been the trend of seafood consumption of each country over the years ? ** - 

Q7 ** What has been the trend of captured vs farmed production of each country over the years ? **   

Q8 ** What has been the trend of sustainable levels of fish stocks in the world ? ** -   

Q9 What is the share of fishes of the that have been been overexploited in the world over the years ? 

Q10 How much of the fish stocks are maintained at sustainable levels in the world of world as compared to overall production level? 

Q11 What is the production level of fish by each continent ?

Q12 What is the consumption level of fish by each continent ?

Q13 What can we learn about the uses of fish catch by countries? 

Q14 What can we learn about the uses of fish catch across the years? 

Q15  Comparing seafood production to seafood consumption across the years 

Q16 What can we learn about the sustainable levels of fishing as compared to farming 

Q17 Would aquaculture alleviate the pressure of seafood consumption across the years? 

Q18 What can we observe about the seafood consumption in coastal countries and landlocked countries over time? 

Q19 What can we observe about the level of seafood being discarded in the world across the years?

Q20 How much fresh water produce in each country over time ? 

# Expected findings 

** What is the contribution of each country in the global fishery sector ? ** 

** What has been the trend of seafood consumption of each country over the years ? ** 

- We would expect that there is an increasing trend of seafood consumption levels over the years given that rise of population. 

- We would expect that coastal countries consumed more seafood than landlocked countries. 

** What has been the trend of captured vs farmed production of each country over the years ? **  

- We would expect that the farmed production to be increasing over the years. 

- We would expect the high technology countries to contribute more to the farmed production. 

** What has been the trend of sustainable levels of fish stocks in the world ? ** 


(dvlp countries > seafood consumed... )... 

# Analysis and findings

[FILL] Should include at least one plot or numerical summary for each of your questions, that helps the reader arrive at an answer. You should also write paragraphs describing the methods, summaries and findings. 

# References
