---
title: "ETC5521 Assignment 1"
subtitle: "Your title"
team: [emu]
author:
  - [FILL]
  - Min Min Soh
date: "`r Sys.Date()`"
output: 
  html_document
---

[This assignment is for ETC5521 Assignment 1 by Team `r rmarkdown::metadata$team` comprising of `r knitr::combine_words(rmarkdown::metadata$author)`.]{style="color:#006DAE;"}

```{r}
library(tidyverse)
```

# Introduction and motivation (Emily)

[FILL] Give the bigger picture of the data, and inspire the reader to learn more about the problem by reading your analysis. 

# Data description 

[FILL] Source of the data, and detail the structure of the data (Rohan)

---- source --- worlddata ... (more sources)
---- table (what kind of data) - a short description of the variable class 
--- 

```{r}
#read data
farmed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/aquaculture-farmed-fish-production.csv')
captured_vs_farmed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/capture-fisheries-vs-aquaculture.csv')
captured <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/capture-fishery-production.csv')
consumption <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/fish-and-seafood-consumption-per-capita.csv')
stock <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/fish-stocks-within-sustainable-levels.csv')
fishery <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/global-fishery-catch-by-sector.csv')
production <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/seafood-and-fish-production-thousand-tonnes.csv')

```

# Questions of interest

[FILL] What questions might be possible to answer with this data, aiming for about 20. Bold the 3-4 questions you will address in the analysis

Q1 What is the contribution of each production sector in global fishery from 1950 ? 

Q2 ** What is the contribution of each country in the global fishery sector ? ** - PICKED  #captured 

Q3 What is the share of type of fishes produced in each country ? 

Q4 What is the production level of each country by capturing over the years ?

Q5 What is the production level of each country by farming over the years  ?

Q6 ** What has beeen the trend of seafood consumption of each country over the years ? ** - PICKED #consumtion dataset 

Q7 ** What has been the trend of captured vs farmed production of each country over the years ? ** - PICKED #captured vs farmed 

Q8 ** What has been the trend of sustainable levels of fish stocks in the world ? ** - PICKED - #stock dataset 

Q9 What is the share of fishes of the that have been been overexploited in the world over the years ? 

Q10 Comparing the sustainable levels of fish stocks of world with the overall production in the world.  

Q11 What is the production level of fish by each continent ?

Q12 What is the consumption level of fish by each continent ?

Q13 What can we learn about the uses of fish catch by countries? 

Q14 What can we learn about the uses of fish catch across the years? 

Q15  Comparing seafood production to seafood consumption across the years 

Q16 What can we learn about the sustainable levels of fishing as compared to farming 

Q17 Would aquaculture alleviate the pressure of seafood consumption across the years? 

Q18 What can we observe about the seafood consumption in coastal countries and landlocked countries over time? 

Q19 What can we observe about the level of seafood being discarded in the world across the years?

Q20 How much fresh water produce in each country over time ? 

# Expected findings (Emily * Rohan)

[FILL] For your chosen questions explain what you think you might find. 

(dvlp countries > seafood consumed... )... 

# Analysis and findings

[FILL] Should include at least one plot or numerical summary for each of your questions, that helps the reader arrive at an answer. You should also write paragraphs describing the methods, summaries and findings. 

# References
