---
title: "ETC5521 Assignment 1"
subtitle: "Global Fishing"
team: [EMU]
author:
  - [Min Min Soh]
  - [Rohan Baghel]
  - [Xiyun Zhou] 
  - [Zhang Zhang]
date: "`r Sys.Date()`"
output:
   bookdown::html_document2:
    toc: yes
    df_print: paged
    number_sections: yes
    fig_caption: yes
editor_options: 
  markdown: 
    wrap: 72
---

[This assignment is for ETC5521 Assignment 1 by Team
`r rmarkdown::metadata$team` comprising of
`r knitr::combine_words(rmarkdown::metadata$author)`.]{style="color:#006DAE;"}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "images/",
                      fig.align = "center")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(tidytuesdayR)
library(plotly)
library(bookdown)
library(visdat)
library(patchwork)
library(maps)
library(kableExtra)
library(readr)
library(ggridges)
library(gganimate)
library(transformr)
library(gifski)
library(viridis)
library(png)
library(GGally)
library(naniar)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

captured_vs_farmed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/capture-fisheries-vs-aquaculture.csv')

consumption <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/fish-and-seafood-consumption-per-capita.csv')

stock <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/fish-stocks-within-sustainable-levels.csv')

production <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/seafood-and-fish-production-thousand-tonnes.csv')

fishery <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-12/global-fishery-catch-by-sector.csv')
```

# Introduction and motivation

Global fishing receives a great deal of attention in the media for the
past decades. The rise of world population has increased the demand for
seafood across the world. Meanwhile, the health of fish population in
the world remains as a concern with the rise of global fishing. Fish
farming (or 'aquaculture')helps to contribute to the seafood production
while alleviating the pressure of wild fisheries.

This study is based on the data sets of global fishing. We explore these
data by analyzing the performance for each country/region in terms of
consumption, production and fish stock level. Moreover, we compare
different data sets to reveal unexpected findings.

We begin by describing the data in the next section, including the way
we source it and prepare it for analysis.

In the analysis section, we present our observations through graphical
displays. Our main tool is R, a programming language for statistical
computing and graphics.

# Data description

The data set has been obtained from `tidytuesday` r package or through
the website <https://ourworldindata.org>. The data comprises of four
files in the ".csv" format which is machine readable and can be used to
analyze the state of fish production and consumption in the world. The
data dictionary for the data set has been given below. They define the
variables and their types in each of the data sets.

## `capture-fisheries-vs-aquaculture.csv`

### 'captured_vs_farmed' - Data Dictionary

Entity: Country/entity Code: Country code Year: Year Aquaculture
production (metric tons): Production of aquaculture Capture fisheries
production (metric tons): Captured aquaculture

```{r}
#str(captured_vs_farmed )
```

The data has 5 variables and 14,674 observations. The 'Code' variable is
character while others are numeric .

### 'captured_vs_farmed' - Data quality Check

```{r}
captured_vs_farmed %>%
  vis_miss(sort_miss = TRUE)
```

The data has 8.1% missing value. Most of them are in the variables of
'aquaculture' and 'code'.

## `fish-and-seafood-consumption-per-capita.csv`

### 'consumption' - Data Dictionary

Entity: Country/entity Code: Country code Year: Year Fish,Seafood-Food
supply quantity(kg/capita/yr)(FAO, 2020):Food supply in fish

```{r}
#str(consumption )
```

As we can see, there are 4 variables and 11,028 observation in total.
The 'Entityy' and 'Code' variable are character while others are
numerical.

### 'consumption' - Data quality Check

```{r}
consumption %>%
  vis_miss(sort_miss = TRUE)
```

15.72% of values in 'Code' are missing. This does not affect our
analysis.

## `fish-stocks-within-sustainable-levels.csv`

### 'stock' - Data Dictionary

Entity: Country/entity Code: code Year: Year Share of fish stocks within
biologically sustainable levels: Share of sustainable fish stock Share
of fish stocks that are overexploited: Share of fish stock that are
overexploited

```{r}
#str(stock)
```

'stock' has 5 variables and 51 observations. The 'Entityy' and 'Code'
variable are character while others are numerical.

### 'stock' - Data quality Check

```{r}
stock %>%
  vis_miss(sort_miss = TRUE)
```

Similar to 'consumption', all missing value are from 'Code'.

## `seafood-and-fish-production-thousand-tonnes.csv`

### 'production' - Data Dictionary

Entity: Country/entity Code: Country code Year: Year\
Pelagic Fish - 2763 - Production - 5510 - tonnes: Pelagic Fish
Crustaceans - 2765 - Production - 5510 - tonnes: Crustaceans
Cephalopods - 2766 - Production - 5510 - tonnes: Cephalopods Demersal
Fish - 2762 - Production - 5510 - tonnes: Demersal Freshwater Fish -
2761 - Production - 5510 - tonnes: Freshwater Molluscs, Other - 2767 -
Production - 5510 - tonnes: Molluscs Marine Fish, Other - 2764 -
Production - 5510 - tonnes: Marine

```{r}
#str(production)
```

There are 10 variables and 10,326 observations. The 'Entityy' and 'Code'
variable are character while others are numerical.

### 'production' - Data quality Check

```{r}
production %>%
  miss_var_summary(sort_miss = TRUE)
```

As we can see, only ' Entity' and 'Code' do not have missing values.

# Expected findings

1.  **What would be the notable finding when we comparing the
    consumption with production?**

-   we expect there is a positive relationship between seafood
    consumption and production

2.  **Compare production of 7 different fish type. Which type of seafood
    contribute the most to meet the increasing demand?**

-   Initial guess: Asia has the highest production of the seafood

3.  **What have been the performances of the top 10 seafood producers?**

-   One of the interesting finding of previous analysis is that seafood
    consumption is highly affected by the geographical factor. Namely,
    coastal. countries tend to consume more than arid regions. Based on
    this, we expect to see the world seafood production is related to
    geographical factor as well. This may implies that the production
    may be concentrated in certain region.

4.  **What has been the trend of fish stocks in the world?**

-   Previous analysis reveals a major decline in world seafood
    consumption in 1989. Here we expect to see that the trend of fish
    stock may show a corresponding movement.

# Analysis and findings

## What would be the notable finding when we comparing the consumption with production?

### Method

Data cleansing: Use `left_join()` function combine `consumption.csv` and
`capture_vs_farmed.csv` into one data set named `con_cap`. In order to
get the total production, add a new variable is the summation of
`Aquaculture production` and `Capture production`. We use the new data
set to do line plots and scatter plot for question one.

### Visualization and analysis

Figure \@ref(fig:worldconsumption) depicts the growing trend in world
seafood consumption from 1969 and 2018. The average yearly seafood
consumption for one person was 13.8kg, while it increased to 20kg in
recent time. Moreover, it is obviously that between 1990 and 1992, the
average seafood consumption decreased, mostly because of economic
problems in low-income nations such those in Africa, Latin America, the
Caribbean, and the Near East. This resulted in increased pressure on
many items' prices (Dumas, 1992). Seafood consumption has grown globally
as the crisis has subsided. Overall, this is consistent with our
assumption that seafood consumption would continue to rise.

The line plot \@ref(fig:production)of world yearly seafood production
shows an upward trend from 1969 to 2018. In 1970, the average production
of seafood is 118 million metric tons. While the production became 3
times greater in recent year. And the production decreasing period
happened same with the seafood consumption. We can make assumption that
there has a positive relationship between seafood consumption and
production.

```{r}
new_con <- consumption %>% 
    filter(Year >= 1969) %>%  
     select(-Code)

con_cap <-new_con %>% 
  left_join(captured_vs_farmed, by = c("Entity","Year")) %>% 
  rename(`seafood consumption(kg/capita/yr)` = `Fish, Seafood- Food supply quantity (kg/capita/yr) (FAO, 2020)`) %>% 
  select(-Code) 

con_cap$`Total production(metric tons)` = rowSums(con_cap[, c("Aquaculture production (metric tons)","Capture fisheries production (metric tons)")])

```

```{r worldconsumption, echo=FALSE, warning = FALSE, message = FALSE, preserve = TRUE, fig.pos = 'H', out.extra = "", fig.cap = "Average seafood consumption in the world over time"}
avg_con <- con_cap %>% 
  group_by(Year) %>%  
  summarise(`average consumption` = mean(`seafood consumption(kg/capita/yr)`),
            `Total production` = sum(`Total production(metric tons)`, na.rm = TRUE), 
            `Total Aquaculture production`=sum(`Aquaculture production (metric tons)`, na.rm = TRUE), 
            `Total Capture fisheries production`=sum(`Capture fisheries production (metric tons)`, na.rm = TRUE) ) %>%  
filter (Year <= 2013)
  

avgcon <- avg_con %>% 
  ggplot(aes(x = Year, y = `average consumption`)) + 
  geom_line(colour = "red") +
  labs(y = "Average Seafood Consumption (kg) per person per year") 

ggplotly(avgcon)

```

```{r production, echo=FALSE, warning = FALSE, message = FALSE, preserve = TRUE, fig.pos = 'H', out.extra = "", fig.cap = "Average yearly seafood production in the world over time"}
totalpro <- avg_con %>%  ggplot(aes(x = Year, y = `Total production`/1000000)) + 
  geom_line(colour = "blue") +
  labs(y = "World Total Production of Seafood(metric tons)") 
  
ggplotly(totalpro)

```

```{r contribution, warning = FALSE, message = FALSE, preserve = TRUE, fig.pos = 'H', out.extra = "", fig.cap = "Contribution of top 10 seafood producer 2015-2018(metric tons)"}

captured_vs_farmed1  <- captured_vs_farmed %>% 
  rename(aquaculture = `Aquaculture production (metric tons)`,
         captured = `Capture fisheries production (metric tons)`)%>% 
  mutate(Entity = toupper(Entity))


captured_vs_farmed_long  <- captured_vs_farmed1 %>% pivot_longer(cols = aquaculture:captured,
                                                           names_to = "production_type",
                                                           values_to = "production")


world_cap_farm <- captured_vs_farmed_long %>% filter(Entity =="WORLD")


world_cap_farmlong <- world_cap_farm %>% pivot_wider(names_from = production_type, values_from= production )%>%
  mutate(total= aquaculture+ captured)%>%
  filter(Year== 2015)
  


dat <- iso3166 %>% 
  mutate(ISOname = toupper(ISOname))  %>%
  rename(Entity = ISOname)

country_cf <- inner_join(captured_vs_farmed1, dat, by = "Entity")

avg_contribution1 <- country_cf %>% 
  filter (Year >= 2015) %>%
  group_by(Entity)%>%
  summarise(avg_captured = mean(captured, na.RM = TRUE),
            avg_aquaculture = mean(aquaculture, na.RM = TRUE))%>%
  arrange(-avg_aquaculture) %>%
  mutate(total= avg_captured+avg_aquaculture)%>%
  arrange(-total) %>%
  mutate(rank = row_number())%>%
  filter(rank %in% 1:10) %>%
  mutate(across(avg_captured:avg_aquaculture, ~ scales::label_comma()(.x))) %>%
  rename(`Average_captured` = avg_captured,
         `Average_aquaculture` = avg_aquaculture)



avg_contribution2015 <-   avg_contribution1 %>%mutate(contribution= total/world_cap_farmlong$total)
   
   
Graph.5g <- ggplot(avg_contribution2015,aes(x=reorder(Entity, (contribution)), y=contribution, fill=Entity))+
  geom_col()+
   theme(axis.text.x = element_text(angle = 60, hjust = 1))+
theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = "Entity", y = "Average production (metric tons)")+
  coord_flip()


ggplotly(Graph.5g)
```

The bar plot \@ref(fig:contribution) shows the contributions of top 10
seafood producer by country. China is the largest contributor and
provide 40% seafood of the whole world in recent years. Indonesia and
India produce 11% and 5.6% respectively. It's intriguing to note that
except US produce 3% and Norway produce 1% of the seafood other major
producers are from Asia.

```{r scatter, echo=FALSE, warning = FALSE, message = FALSE, preserve = TRUE, fig.pos = 'H', out.extra = "", fig.cap = "Relationship between seafood consumption and production"}
ggscatmat(avg_con, 
          columns = 2:ncol(avg_con)) 
```

To explore relationship between seafood consumption and seafood
production, we made a scatter matrix plot \@ref(fig:scatter). Since
seafood produce in two ways: aquaculture and capture, we add another two
variables `Aquaculture production` and `Capture production`to do the
scatter plot and try to find some interest between each variable. The
outcome demonstrates a strong positive correlation between each
variable. There is a significant linear positive relation between
average consumption and total production as we expected.The correlation
coefficient is 0.97 almost reach to 1. Moreover, the two methods used to
produce fish also has a positive linear relation.

### Summary and interesting findings

1.  Upward trend for seafood production and consumption worldwide\
2.  Positive linear relationship between seafood production and
    consumption.
3.  Two seafood produce method: Aquaculture and capture have positive
    linear relation.
4.  Asia is the major continent to produce seafood and China produces
    the most seafood.

## Compare production of 7 different fish type. Which type of seafood contribute the most to meet the increasing demand?

### Method

We want to explore something interest when comparing fish type
production.

Use the `contribution.csv` data set and simplified the name of the
variables of 7 types of fish, since they are too long and not easy to
read. Add up each fish type as a total production variable. We wonder
which country contribute the most for each type of fish. In order to to
do that, I make a summary for average yearly production for each type of
fish by countries and then using max() function to find the answer.

To compare production of each type fish, I decide to use column plot.
Not only it can shows the trend but also easy to see the difference.
Categorize year to 4 decade and get the average production for each type
fish. Then use `pivot_longer()` to gather fish type into one column,
this is the preparation for making the column plot.

```{r}

new_pro<-production %>% 
  rename(`Pelagic Fish`=`Commodity Balances - Livestock and Fish Primary Equivalent - Pelagic Fish - 2763 - Production - 5510 - tonnes`, 
         Crustaceans = `Commodity Balances - Livestock and Fish Primary Equivalent - Crustaceans - 2765 - Production - 5510 - tonnes`, 
         Cephalopods = `Commodity Balances - Livestock and Fish Primary Equivalent - Cephalopods - 2766 - Production - 5510 - tonnes`, 
        `Demersal Fish`  = `Commodity Balances - Livestock and Fish Primary Equivalent - Demersal Fish - 2762 - Production - 5510 - tonnes`, 
        `Freshwater Fish` = `Commodity Balances - Livestock and Fish Primary Equivalent - Freshwater Fish - 2761 - Production - 5510 - tonnes`, 
          Molluscs = `Commodity Balances - Livestock and Fish Primary Equivalent - Molluscs, Other - 2767 - Production - 5510 - tonnes`, 
       `Marine Fish` = `Commodity Balances - Livestock and Fish Primary Equivalent - Marine Fish, Other - 2764 - Production - 5510 - tonnes`)


new_pro$`Total fish production`  = rowSums(new_pro[, c("Pelagic Fish", "Crustaceans","Cephalopods", "Demersal Fish", "Freshwater Fish", "Molluscs", "Marine Fish")])

```

```{r}
fish_type1 <- new_pro %>%  
  group_by(Entity) %>% 
  summarise(`avg Pelagic Fish` = mean(`Pelagic Fish`, na.rm= TRUE), 
            `avg Crustaceans` = mean(Crustaceans,  na.rm= TRUE), 
            `avg Cephalopods` = mean(Cephalopods,  na.rm= TRUE), 
            `avg Demersal Fish` = mean(`Demersal Fish`,  na.rm= TRUE), 
            `avg Freshwater Fish` = mean(`Freshwater Fish`,  na.rm= TRUE), 
            `avg Molluscs` = mean(Molluscs,  na.rm= TRUE),
            `avg Marine Fish` = mean(`Marine Fish`,  na.rm= TRUE)) %>%  
  na.omit() 

fish_type<-fish_type1[-c(126),]

#max(fish_type$`avg Marine Fish`)

#fish_type %>% filter(`avg Marine Fish` >= 6597212)

```

```{r fishtype, echo=FALSE, warning = FALSE, message = FALSE, preserve = TRUE, fig.pos = 'H', out.extra = "", fig.cap = "Compare Fish Type production"}

yr_fishtype <- within(new_pro, {
  Decade <- NA
  Decade[Year>=1970 & Year <1980] <- "1970-1980"
  Decade[Year >=1980 & Year <1990] <- "1980-1990"
  Decade[Year>=1990 & Year <2000] <- "1990-2000"
  Decade[Year >=2000 & Year <2013] <- "2000-2013"
})

dec_type <- yr_fishtype %>%  
  group_by(Decade) %>% 
  summarise(`PelagicFish` = mean(`Pelagic Fish`, na.rm= TRUE), 
            `Crustaceans` = mean(Crustaceans,  na.rm= TRUE), 
            `Cephalopods` = mean(Cephalopods,  na.rm= TRUE), 
            `DemersalFish` = mean(`Demersal Fish`,  na.rm= TRUE), 
            `FreshwaterFish` = mean(`Freshwater Fish`,  na.rm= TRUE), 
            `Molluscs` = mean(Molluscs,  na.rm= TRUE),
            `MarineFish` = mean(`Marine Fish`,  na.rm= TRUE)) %>%  na.omit()

col_type_fish <- dec_type%>% pivot_longer(cols=c('PelagicFish', 'Crustaceans', 'Cephalopods','DemersalFish','FreshwaterFish','Molluscs','MarineFish'),
                    names_to='Fish Type',
                    values_to='production(tons)') 


ggplot(col_type_fish, aes(`Fish Type`,`production(tons)`)) +
      geom_col(aes(fill = `Fish Type`)) +  
        facet_grid(. ~Decade) +
  theme(axis.text.x = element_text(angle=45, hjust = 1))

#str(col_type_fish)


```

### Visualization and analysis

From the column plot \@ref(fig:fishtype) we can see that the Pelagic Fish
has the largest production scale when comparing with other type fish.
Demersal Fish used to take the second largest contribution of fish
product. However, the second largest changed to Freshwater Fish in
recent years. The overall trend for each type fish is increasing along
with time. While the production of Pelagic Fish decreased in this decade
and Demersal Fish almost keep a same level.

The table below shows the maximum yearly production of 7 different fish
type by countries. Americas has the largest production for Pelagic Fish.
Asia contribute most for Crustaceans, Cephalopods, Freshwater Fish,
Molluscs and Marine Fish. And European countries contribute most of the
Demersal Fish.

#### `Maximum yearly production of 7 different fish type`

| Fish Type       | Entity   | production(metric tons) |
|:----------------|:---------|:------------------------|
| Pelagic Fish    | Americas | 12495788                |
| Crustaceans     | Asia     | 3748006                 |
| Cephalopods     | Asia     | 1578872                 |
| Demersal Fish   | Europe   | 7979733                 |
| Freshwater Fish | Asia     | 14040488                |
| Molluscs        | Asia     | 6115966                 |
| Marine Fish     | Asia     | 6597212                 |

### Summary and interesting findings

1.  Pelagic fish takes the largest fish production.
2.  Asia contribute the most for fish production.

## What have been the performances of the top 10 seafood producers?

### Method:

We are going to specify the analysis area by focusing on the top 10
seafood producers and digging the data from a wider prospective.

Here we use 'captured_vs_farmed' data set. The first reason is that
'captured_vs_farmed' contains less NA value than 'production'. Secondly,
we tend to focus on the trend of 'captured vs farmed' seafood production
for those top producers.

In addition, the use of sorted data from the original report help us
quite a lot to simplify the process. These data includes: 'country_cf',
'dat', 'world_cap_farm' and 'captured_vs_farmed_long'.

To answer this question, we took the following steps:

1 We worked out the overall top ten producers between 2015-2018 as well
as the top 10 in terms of captured or farmed fish during the same
period. We took the average production for each country 2015 - 2018
because there are many missing value in 2018. This allows us to present
a accurate picture to reflect the most recent seafood productivity of
each country.

2 Next we figured out the top ten producers of 1960s using data between
1960-1970 due to the problem of missing value.Then we compared the top
10 of 1960s with the list of 2010s.

3 We would take a look at the percentage contribution for each of the
top producers.

4 Lastly, we tried to find out what were the changes of production
between 1960s and 2010s.

### Visualization and analysis

```{r}
captured_vs_farmed  <- captured_vs_farmed %>% 
         rename(aquaculture = `Aquaculture production (metric tons)`,
         captured = `Capture fisheries production (metric tons)`)
captured_vs_farmed<- captured_vs_farmed %>% 
         mutate(Entity = toupper(Entity))
dat <- iso3166 
dat <- dat %>% mutate(ISOname = toupper(ISOname)) 
dat <-  dat %>%rename(Entity = ISOname)
consumption <-  consumption %>% mutate(Entity = toupper(Entity))
country_cf <- inner_join(captured_vs_farmed, dat, by = "Entity")
captured_vs_farmed_long  <- captured_vs_farmed %>% 
        pivot_longer(cols = aquaculture:captured,names_to = "production_type",values_to = "production")
avg_contribution1 <- country_cf %>% 
        filter (Year >= 2015) %>%
        group_by(Entity)%>%
        summarise(avg_captured = mean(captured, na.RM = TRUE),
            avg_aquaculture = mean(aquaculture, na.RM = TRUE))%>%
        arrange(-avg_aquaculture) %>%
        mutate(total= avg_captured+avg_aquaculture)%>%
        arrange(-total) %>%
        mutate(rank = row_number())%>%
        filter(rank %in% 1:10) %>%
        mutate(across(avg_captured:avg_aquaculture, ~ scales::label_comma()(.x))) %>%
        rename(`Average_captured` = avg_captured,
               `Average_aquaculture` =    avg_aquaculture)
avg_aquaculture1 <- country_cf %>% 
        filter (Year >= 2015) %>%
        group_by(Entity)%>%
        summarise(avg_aquaculture = mean(aquaculture, na.RM = TRUE))%>%
        arrange(-avg_aquaculture) %>%
        mutate(rank = row_number())%>%
        filter(rank %in% 1:10) %>%
        rename(`avg_aquaculture` = avg_aquaculture)
avg_captured1 <- country_cf %>% 
        filter (Year >= 2015) %>%
        group_by(Entity)%>%
        summarise(avg_captured = mean(captured, na.RM = TRUE))%>%
        arrange(-avg_captured) %>%
        mutate(rank = row_number())%>%
        filter(rank %in% 1:10) %>%
        rename(`avg_captured` = avg_captured)
```



```{r}


Graph.5a <- ggplot(avg_contribution1, aes(x= reorder(Entity, -total), y=total/1000000, fill=Entity)) + 
        geom_col( )+
        scale_fill_brewer(palette = "Set3") +
        theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
        labs(x = "Entity", y = "production ( millions metric tons)")+
        ggtitle("Graph.5a: Tope 10 producers 2010s")
Graph.5b <- ggplot(avg_aquaculture1, aes(x= reorder(Entity, desc(avg_aquaculture)),  y=avg_aquaculture/1000000,fill=Entity)) + 
        geom_col( )+
        scale_fill_brewer(palette = "Set3") +
        theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
        labs(x = "Entity", y = " production (millions metric tons)")+
        ggtitle("Graph.5b: Tope 10 aquaculture producers 2010s ")
Graph.5c <- ggplot(avg_captured1, aes(x= reorder(Entity, desc(avg_captured)), y=avg_captured/1000000,fill=Entity)) + 
        geom_col( )+
        scale_fill_brewer(palette = "Set3") +
        theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
        labs(x = "Entity", y = " production (millions metric tons)")+
        ggtitle("Graph.5c: Top 10 captured producers 2010s ")
Graph.5a 
Graph.5b  
Graph.5c
```

From the 3 plots above, we can see that China
and Indonesia are the top 2 producers in each of the three ranks. China
takes the dominate position of seafood production.

In addition, the U.s and Peru are the overall top producer but did not
shown on the rank of aquaculture fishing. Similarly, Philippines and
Bangladesh are not listed in the top captured seafood producer. To sum
up, only six countries present a strong productivity in both captured
and farmed production.

```{r}
contribution70 <- country_cf %>% 
  filter (Year<=1970) %>%
  group_by(Entity)%>%
  mutate_all(~replace(., is.na(.), 0))%>%
  summarise(avg_captured = mean(captured, na.RM = TRUE),
            avg_aquaculture = mean(aquaculture, na.RM = TRUE))%>%
  mutate(total= avg_captured+avg_aquaculture)%>%
  arrange(-total) %>%
  mutate(rank = row_number())%>%
  filter(rank %in% 1:10) %>%
  mutate(across(avg_captured:avg_aquaculture, ~ scales::label_comma()(.x))) 
aquaculture70 <- country_cf %>% 
  filter (Year <= 1970) %>%
  group_by(Entity)%>%
  summarise(avg_aquaculture = mean(aquaculture, na.RM = TRUE))%>%
  arrange(-avg_aquaculture) %>%
  mutate(rank = row_number())%>%
  filter(rank %in% 1:10) %>%
  rename(`avg_aquaculture` = avg_aquaculture)
captured70 <- country_cf %>% 
  filter (Year <= 1970) %>%
  group_by(Entity)%>%
  summarise(avg_captured = mean(captured, na.RM = TRUE))%>%
  arrange(-avg_captured) %>%
  mutate(rank = row_number())%>%
  filter(rank %in% 1:10) %>%
  rename(`avg_captured` = avg_captured)
Graph.5d <- ggplot(contribution70, aes(x= reorder(Entity, -total), y=total/1000000,fill=Entity)) + 
  geom_col( )+
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = "Entity", y = "production (millions metric tons)")+
  ggtitle("Graph.5d: Tope 10 producers 1960s")
Graph.5e <- ggplot(aquaculture70, aes(x= reorder(Entity, desc(avg_aquaculture)),y=avg_aquaculture/1000000,fill=Entity )) + 
  geom_col( )+
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = "Entity", y = "production (millions metric tons)")+
  ggtitle("Graph.5e: Tope 10 aquaculture producer 1960s")
Graph.5f <- ggplot(captured70, aes(x= reorder(Entity, desc(avg_captured)), y=avg_captured/1000000, fill=Entity)) + 
  geom_col( )+
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = "Entity", y = " ")+
  ggtitle("Graph.5f: Top 10 captured producer 1960s")
```

```{r}
Graph.5d +Graph.5a 
```

Compared the top 10 producer in 1960s with that of 2010s, we can see
Graph.5d +Graph.5a that three developed countries:South Africa, Canada and
Spain were the top 10 in 1960s but did not shown on the list of 2010s.On
the other hand, three developing countries: Vietnam, Philippines and
Bangladesh are the new countries listed in 2010s rank. Geographically,
all the new comers in 2010s are located in South-east Asia.

Moreover, considering what we observed in Graph.5a-5c, it seems like the
high aquaculture production is the main reason for Vietnam, Philippines
and Bangladesh to be ranked as top 10 producer in 2010s.

```{r change, echo=FALSE, warning = FALSE, message = FALSE, preserve = TRUE, fig.pos = 'H', out.extra = "", fig.cap =""}
 

avg_contribution1970 <- country_cf %>% 
       filter (Year<=1970) %>%
       group_by(Entity)%>%
       summarise(avg_captured = mean(captured, na.RM = TRUE),
       avg_aquaculture = mean(aquaculture, na.RM = TRUE))%>%
       mutate_all(~replace(., is.na(.), 0))%>%
       mutate(total1970= avg_captured+avg_aquaculture)%>%
       filter(Entity %in% c("CHINA","INDONESIA", "INDIA", "VIETNAM", "UNITED STATES", "PERU","JAPAN", "PHILIPPINES", "BANGLADESH", "NORWAY"))%>%
       mutate(across(avg_captured:avg_aquaculture, ~ scales::label_comma()(.x))) 
Top10_percentage_change <- avg_contribution1970 %>%
       dplyr::select(Entity, total1970)%>%
       full_join(avg_contribution1, by=NULL)%>%
       rename("total2015"=total)%>%
       dplyr::select(Entity, total1970,total2015)%>%
       mutate(percentage_change = (total2015-total1970)/total1970*100)%>%
       arrange(desc(percentage_change))
Graph.5h <- ggplot(Top10_percentage_change, aes(x=reorder(Entity, percentage_change), y=percentage_change))+
geom_col(col = ifelse(Top10_percentage_change$percentage_change < 0,'red','green') )+
      theme(axis.text.x = element_text(angle = 60, hjust = 1))+
      labs(x = "Entity", y = "Percentage change 1960s-2010s")+
      ggtitle("Graph.5h: Percentage change of top 10 producers 1960s-2010s")+
      coord_flip()

 ggplotly(Graph.5h)
```

It can be seen from Graph.5h that both China and Indonesia
showed a more than 20 times increase in seafood production from
1960-2018. Following this are the tier 2 countries: Vietnam, India,
Bangladesh and Philippines, a huge increase between 5 to 10 times. The
U.s and Norway showed a 110% and 59% rise respectively.

On the other hand, Peru and Japan experienced a considerable decline in
seafood production, which were 37% and 41%. However, they are still the
top 10 producer all over the world.

```{r cap, echo=FALSE, warning = FALSE, message = FALSE, preserve = TRUE, fig.pos = 'H', out.extra = ""}
cf_data_long1 <- captured_vs_farmed_long  %>% 
       filter (Year >= 1960) %>%
       mutate(Entity = toupper(Entity)) %>%
       filter (Entity %in% c(avg_contribution1$Entity))
Graph.5i <- cf_data_long1%>%
       ggplot(aes(x = Year,y = (production/1000000),colour = production_type))+
       geom_line()+
       labs(x = "Year", y = "Production (metric tons in millions)")+
       scale_y_continuous(labels = scales::label_comma(scale = 1e-0))+
       scale_x_continuous(breaks = seq(1960, 2017, by= 10))+ 
       facet_wrap(~ Entity, scales =  "free", nrow = 3)+
       theme(axis.text.x = element_text(angle = 60, hjust = 1))+
       scale_color_viridis(discrete = TRUE)+
       ggtitle("Graph.5i: Captured vs aquaculture  1960-2018")
ggplotly(Graph.5i)
```

Here we tend to further explore what happened to Japan and Peru. As we
can see from Graph.5i, the farmed production of these two countries
remains at the same level from 1960-2018, while the captured production
declined quite a lot compare to the peak.

Moreover, compared Graph.5i and Graph.5h, it is clearly
that the countries with relatively lower increase (The U.s, Norway) or
decline (Peru, Japan) in the total production, all presenting a poor
development in farmed production. In other words, those four countries
produce more captured fish than farmed seafood.

In comparison, other six countries with relatively higher overall
increase, experiencing a significant upward trend in farmed production.
The level of aquaculture fishing of these six countries are higher than
captured fish.

### Summary and interesting findings:

1 Developing countries take more position in top ten seafood producers
compared with 1960s..

2 ALL the new comers on the top 10 list in 2010s are from South-east
Asia.

3 The center of world's seafood production is now in West pacific and
Southeast Pacific.

4 Farmed production contribute more than captured seafood to the
improvement of countries' overall productivity.

5 China and Indonesia are the top 2 seafood producers in total
production as well as the farmed and captured ranking. Over the last 60
years, these two countries experienced the most significant upward trend
in seafood production.

6 Peru and Japan were the top 2 countries in 1960s. However, they were
the only two countries which showed a decrease in seafood production
within the top 10 list over the last 60 years.

## What has been the trend of fish stocks in the world?

### Method:

Here we use score 'stock' and 'consumption' data. We not only analysed
data within 'stock', but also compare 'stock' with 'consumption'.

Similarly, we used some cleaned data from original report including
'consumption' and 'annual_consumption'.

1 We firstly plotted a graph to have a comparison of sustainable vs
overexploited based on trend and proportion.

2 Previous analysis reveals the world's trend based on the stock type.
Here we would like to go further by analyzing the trend based on each
region.

3 Also, we are going to compare the world's trend of overexploited stock
with the trend of world seafood consumption. A considerable decline in
seafood consumption around 1989 was observed in previous analysis. By
comparing these two data sets, we would like to see if any similar
pattern exist.

### Visualization and analysis

```{r}
stock_world <- stock %>% 
      mutate(Sustainable_levels_Fish = 
      stock$`Share of fish stocks within biologically sustainable levels (FAO, 2020)`,
      Overexploited_Fish = stock$`Share of fish stocks that are overexploited`) %>%
      dplyr::select(-`Share of fish stocks within biologically sustainable levels (FAO, 2020)`, -`Share of fish stocks that are overexploited`) %>%
      slice(31:51) %>% 
      dplyr::select(-Code)
stock_worldL <- stock_world %>% 
      pivot_longer(cols = Sustainable_levels_Fish	:Overexploited_Fish,names_to = "type", values_to = "proportion") 
Graph.6a <- ggplot(stock_worldL) +
geom_col(aes(x = Year, y = proportion, fill = type, position = "fill",width = 0.8)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1))+
      labs(x = "Year", y = "Percentage to the world ")+
      scale_x_continuous(breaks = seq(1974, 2017, by= 2))+
      ggtitle("Graph.5h: Percentage change of stock level")
      
ggplotly(Graph.6a)
```

It is obviously that sustainable stock is always having a large
proportion then those overexploited.

Although the level of overexploited stock increasing with fluctuation
overtime, it has never exceeded 40% of the total.

```{r}
stock_worldL1 <- stock_worldL %>% filter(type=='Overexploited_Fish')
Graph.6b<- ggplot( stock_worldL1, aes(x=Year, y=proportion))+
      geom_line(colour = "red")+
      scale_x_continuous(breaks = seq(1974, 2017, by= 4))+
      labs(x = "Year", y = "Proportion of Overexploited Fish")+
      ggtitle("Graph.6b: Trend of Overexploited Fish % ")+
      geom_vline(xintercept = 1989,  color = "blue", size=0.5)+
      geom_vline(xintercept = 1990,  color = "green", size=0.5)+
      geom_vline(xintercept = 1995,  color = "red", size=0.5)+
      geom_vline(xintercept = 1997,  color = "pink", size=0.5)+
      geom_vline(xintercept = 2000,  color = "yellow", size=0.5)+
      geom_vline(xintercept = 2004,  color = "purple", size=0.5)+
      geom_vline(xintercept = 2008,  color = "black", size=0.5)+
      geom_vline(xintercept = 2011,  color = "orange", size=0.5)
consumption <- consumption %>% 
      rename(consumption = `Fish, Seafood- Food supply quantity (kg/capita/yr) (FAO, 2020)`)
annual_consumption <- consumption %>% 
      group_by(Year) %>%
      summarise(av_consumption = mean(consumption, na.rm=TRUE))
annual_consumption1<- annual_consumption %>% 
      filter(Year>'1973')
Graph.6c <-ggplot(data = annual_consumption1,
         aes(x = Year, y = av_consumption)) +
      geom_line(colour = "blue") +
      labs(y = "Average consumption") +
      scale_x_continuous(breaks = seq(1974, 2017, by= 3))+
      ggtitle("Graph.6c: Trend of Seafood consumption ")+
      geom_vline(xintercept = 1989,  color = "blue", size=0.5)+
      geom_vline(xintercept = 1992,  color = "green", size=0.5)+
      geom_vline(xintercept = 1997,  color = "red", size=0.5)+
      geom_vline(xintercept = 1998,  color = "pink", size=0.5)+
      geom_vline(xintercept = 2001,  color = "yellow", size=0.5)+
      geom_vline(xintercept = 2006,  color = "purple", size=0.5)+
      geom_vline(xintercept = 2011,  color = "black", size=0.5)+
      geom_vline(xintercept = 2013,  color = "orange", size=0.5)
Graph.6b /Graph.6c
```

```{r}
consumption_stock <- left_join(stock_world, annual_consumption1, by='Year')%>%
        rename('Consumption_(kg/capita/yr)'= av_consumption)
Graph.6d<- ggpairs(consumption_stock, columns = 3:5)
Graph.6d
```

Here we compared the proportion of overexploited stock with the world
seafood consumption.

We plotted 8 vertical lines.Each of them represents a start / end point
of a particular up / downward trend. Also, vertical lines with the same
color corresponds to each other in different graphs. Since a major
decline in consumption in 1989 was observed in previous analysis, we set
up 1989 as the first spot.

It can be seen from  Graph.6b /Graph.6c that the trend of overexploited
stock and world seafood consumption have a similar pattern. For each
individual trend between two lines in Graph.6a, we can find out a
similar trend between two lines with the corresponded color in Graph.6b.

It is interesting that most of the line (except the trend between yellow
and purple) in the Graph.6b are on th left to the line with same color
in Graph.6c. This implies that the trend in consumption always delayed
compared with the corresponding trend in overexploited stock.

Moreover, Graph.6d indicates that sustainable stock having a strong negative
correlation with seafood consumption, which further proved our
observation in Graph.6b and 6c. 

```{r 6f, echo=FALSE, warning = FALSE, message = FALSE, preserve = TRUE, fig.pos = 'H', out.extra = ""}
stockL <- stock %>% 
       pivot_longer(cols = 'Share of fish stocks within biologically sustainable levels (FAO, 2020)':'Share of fish stocks that are overexploited', names_to= "Type", values_to = "Proportion")  %>% 
       filter(Entity!= "World") 
stockL_sustainable <- stockL %>% 
       filter(Type=='Share of fish stocks within biologically sustainable levels (FAO, 2020)')
stockL_overexploited <- stockL %>% 
       filter(Type=='Share of fish stocks that are overexploited')
  Graph.6d <- ggplot(stockL_sustainable, aes(x = Entity, y = Proportion , fill = Entity)) +
       geom_col(position = "stack") +
       facet_wrap(Year~ Type, nrow=2)+
       theme(axis.text.x = element_text(angle = 60,  hjust = 1)) 
  
  
  Graph.6e <- ggplot(stockL_overexploited, aes(x = Year, y = Proportion , colour = Entity)) +
       geom_line() +
       facet_wrap( ~ Entity, nrow=3)+
       theme(axis.text.x = element_text(angle = 60,  hjust = 1)) 
  
  
 
  stockL_sustainableW  <- stock %>% 
      pivot_longer(cols = 'Share of fish stocks within biologically sustainable levels (FAO, 2020)':'Share of fish stocks that are overexploited', names_to= "Type", values_to = "Proportion") %>%
      filter(Type=='Share of fish stocks within biologically sustainable levels (FAO, 2020)')%>% 
      filter(Year %in% c("2015","2017"))%>%
      dplyr::select(-Type )%>%
      dplyr::select(-Code )%>%
      pivot_wider(names_from = Year, values_from = Proportion)%>%
      rename('Sustainable2017'='2017','Sustainable2015'='2015')%>%        mutate(percentage_change= (Sustainable2017-Sustainable2015)/Sustainable2015*100)%>%
      mutate_if(is.numeric, round, digits = 4)  
  
    
   Graph.6f <- ggplot(stockL_sustainableW,aes(x=reorder(Entity, percentage_change), y=percentage_change, fill=Entity))+
      geom_col(col = ifelse(stockL_sustainableW$percentage_change == !0,'red','green'))+
      theme(axis.text.x = element_text(angle = 60, hjust = 1))+
      labs(x = "Entity", y = "Percentage change ")+
      ggtitle("Graph.6f: Change of sustainable stock level 2015-17")+
      coord_flip()
     
     
ggplotly(Graph.6f)
  
```

Next, we plotted graphs to check the stock level based on regions.

As we can see from Graph.6f,the world's sustainable stock declined
1.2% between 2015-2017. The main factor contributed to this drop is the
decreased sustainable stock in Northwest Pacific and Northwest Central
Atlantic which were 21% and 24% respectively.

In comparison, Southeast Pacific, Southwest Pacific and Southwest
Atlantic experienced the largest increase in sustainable stock.

It is worth noting that the stock level of Southeast Central
Atlantic,Eastern Central Pacific and Eastern Central Atlantic were
almost unchanged.

### Summary and interesting findings:

1 The level of overexploited stock increasing with fluctuation overtime.

2 There are similar pattern between the trend of overexploited stock and
world seafood consumption since 1989.

3 Compared with stock level, there are lags shown in the world
consumption for each similar trend.

4 Some regions indicate very stable stock level between 2015-17.

# Conclusion

The seafood consumption and production keep increase with the time
change. Moreover, it is noteable that a strongly positive linear
relationship between fish consumption and production. China produces the
most seafood in worldwide and Asia is the main continent producing
seafood. Pelagic fish has the highest production scale and most of them
comes from Americas.

Compared with 1960s, developing countries from South-east Asia take more
seats in top 10 seafood producers in 2010s. Also, farmed fishing is the
most important factor to the improvement overall productivity. Moreover,
we observe that overexploited stock has strong positive correlation with
world fish consumption.

# Limitation and recommendation

1 With missing value and incomplete data in 'production' and 'captured
vs farmed', this study may lack accuracy for visualization.

2 Only 2 years data is available for most entities in 'stock' data set,
which make it difficult to analyzing long term performance.

3 There is not a clear definition about those 'Entity' in 'stock' data.
In this case, we are unable to link those entities with countries to
have a deeper analysis.

4 Future research would benefit from the use of more datasets to broaden
the prospective of analysis. For example, a data of stock level based on
each country's performance may help us to dig more insights in terms of
sustainable fishing.

# References

Becker OScbRA, Minka ARWRvbRBEbTP, Deckmyn. A (2021). *maps: Draw
Geographical Maps*. R package version 3.4.0,
<https://CRAN.R-project.org/package=maps>.

C. Sievert. Interactive Web-Based Data Visualization with R, plotly, and
shiny. Chapman and Hall/CRC Florida, 2020.

Dumas, M. W. (1992). Productivity trends: prepared fish and seafoods
industry. Monthly Lab. Rev., 115, 3.

Hughes E (2022). *tidytuesdayR: Access the Weekly 'TidyTuesday' Project
Dataset*. R package version 1.0.2,
<https://CRAN.R-project.org/package=tidytuesdayR>.

H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag
New York, 2016.

Ooms J (2022). *gifski: Highest Quality GIF Encoder*. R package version
1.6.6-1, <https://CRAN.R-project.org/package=gifski>.

Pedersen T, Robinson D (2020). *gganimate: A Grammar of Animated
Graphics*. R package version 1.0.7,
<https://CRAN.R-project.org/package=gganimate>.

Pedersen T (2020). *patchwork: The Composer of Plots*. R package version
1.1.1, <https://CRAN.R-project.org/package=patchwork>.

Pedersen T (2020). *transformr: Polygon and Path Transformations*. R
package version 0.1.3, <https://CRAN.R-project.org/package=transformr>.

Simon Garnier, Noam Ross, Robert Rudis, Antônio P. Camargo, Marco
Sciaini, and Cédric Scherer (2021). Rvision - Colorblind-Friendly Color
Maps for R. R package version 0.6.2.

Simon Garnier, Noam Ross, Robert Rudis, Antônio P. Camargo, Marco
Sciaini, and Cédric Scherer (2021). Rvision - Colorblind-Friendly Color
Maps for R.

Subasinghe, R., Soto, D., & Jia, J. (2009). Global aquaculture and its
role in sustainable development. Reviews in aquaculture, 1(1), 2-9.

Tierney N (2017). "visdat: Visualising Whole Data Frames." *JOSS*,
*2*(16), 355. <doi:10.21105/joss.00355>
<https://doi.org/10.21105/joss.00355>,
<http://dx.doi.org/10.21105/joss.00355>.

Urbanek S (2013). *png: Read and write PNG images*. R package version
0.1-7, <https://CRAN.R-project.org/package=png>.

Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R,
Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E,
Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K,
Vaughan D, Wilke C, Woo K, Yutani H (2019). "Welcome to the tidyverse."
*Journal of Open Source Software*, *4*(43), 1686.
<doi:10.21105/joss.01686> <https://doi.org/10.21105/joss.01686>.

Wickham H, Hester J, Bryan J (2022). *readr: Read Rectangular Text
Data*. R package version 2.1.2,
<https://CRAN.R-project.org/package=readr>.

Wilke C (2021). *ggridges: Ridgeline Plots in 'ggplot2'*. R package
version 0.5.3, <https://CRAN.R-project.org/package=ggridges>.

Yihui Xie (2022). bookdown: Authoring Books and Technical Documents with
R Markdown. R package version 0.26.

Zhu H (2021). *kableExtra: Construct Complex Table with 'kable' and Pipe
Syntax*. R package version 1.3.4,
<https://CRAN.R-project.org/package=kableExtra>.



